A Switching Planner for Combined Task and Observation Planning

### 2010-11-26 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/switching.tex
changed the related text to fix error

We describe our continual planning approach that switches between
sequential and contingent sessions. Continual planning and execution
proceeds in a deterministic-action POMDP described in DTPDDL. During a
sequential session, planning is performed by a ``classical''
system,\footnote{That is, a system designed to solve fully observable
deterministic tasks.}  and execution proceeds according to the {\em
trace} computed by that system. Taking the form of a classical plan,
the {\em trace} specifies a sequence of POMDP actions that achieve the
agent's objectives in a deterministic approximation, i.e., {\em
determinisation}, of the problem at hand. More precisely, the trace is
an interleaved sequence of POMDP actions and {\em assumptive}
actions. The latter correspond to assumptions the planner makes about
the truth value of {\em runtime state variables} -- e.g. that a box of
cornflakes is located in the kitchen at the third plan step. Another
kind of assumption, called an {\em applicability} assumption, is made
when the trace includes an action $\action$ that is not applicable
with probability $1$ at the belief-state
\bstate\ that the system is projected to be in when \action\ is executed, 
i.e., $\exists\state\in\states\;
\bstate(\state) > 0$ and $\poss(\action)\not\subseteq\state$. By
scheduling
\action,  the serial planner makes an assumption about the
observability of the precondition $\poss(\action)$.

Our approach always begins with a sequential session. Non-assumptive
actions from the trace are executed until the applicability of the
next scheduled action, here \action, is questionable. A contingent
session then begins, which tailors sensory processing by the agent to
perceive whether the assumptions made in the trace hold. For each {\em
active} assumption in the trace, we add a {\em disconfirm} action to
the POMDP whose execution is rewarding (resp. costly) if the
assumption is false (resp. true). We also add one {\em confirm} action
that is rewarding (resp. costly) to execute if $\poss(\action)$ is
true (resp. false). If execution of the contingent plan applies a
disconfirm action, then a new sequential session begins at the current
underlying belief-state. If the confirm action is executed, the
sequential session is resumed, and \action\ is executed. Otherwise,
control rests in the continual session.

Because contingent planning is only practical in relatively small
POMDPs, contingent sessions proceed in an abstract decision process
determined by the current trace. That abstract process is constructed
by first omitting all variables from the model that are not featured
in the trace, and by then iteratively refining that abstraction while
the resulting model is of a practicable size.

@@

For sequential planning the operators from the POMDP model are
available with semantics that accommodate the above abstraction as
follows. In the deterministic model a proposition $\prop$ can be
thought to have a ternary interpretation at a state \state, as either
{\em true}, written $\prop\in\state$, {\em false}, written
$\prop\not\in\state$, or {\em unspecified}, with a slight abuse of
notation written $\prop\#\in\state$. For example, in $\pp{DORA}$
$(=(\pp{is-in}~\pp{box})\pp{office})\#\in\state_0$. For POMDP actions
$\action\in\actions$ in the deterministic model, if
$\prop \in \poss(\action)$, or if $\prop$ is the subject of a positive
or negative effect of $\action$, then $\action$ is not applicable
in \state\ if $\prop\#\in\state$. Indeed, first an assumptive action
must be applied to make a truth assignment to \prop, and then that
truth assignment could be modified by POMDP operators. Lastly, we
address the switching semantics of action execution. If
$\action\in\actions$ has a precondition that is not true with
probability $1$ in the underlying belief $\bstate$, then before it is
executed blindly, the system switches to a contingent session,
discussed in detail in the following section.

>>>>

We describe our continual planning approach that switches between
sequential and contingent sessions. This proceeds by interleaved
planning and execution in a deterministic-action POMDP described in
DTPDDL. During a sequential session, planning is performed by a
``classical'' system,\footnote{That is, a system designed to solve
fully observable deterministic tasks.}  and execution proceeds
according to the {\em trace} computed by that system. Taking the form
of a classical plan, the trace specifies a sequence of POMDP
actions that achieve the agent's objectives in a deterministic
approximation, i.e., {\em determinisation}, of the problem at
hand. More precisely, the trace is an interleaved sequence of POMDP
actions and {\em assumptive} actions. The latter correspond to
assumptions the planner makes about the truth value of {\em runtime
state variables} -- e.g. that a box of cornflakes is located in the
kitchen at the third plan step. Another kind of assumption, called an
{\em applicability} assumption, is made when the trace includes an
action $\action$ that is not applicable with probability $1$ at the
belief-state
\bstate\ that the system is projected to be in when \action\ is
scheduled for execution, i.e., $\exists\state\in\states\;
\bstate(\state) > 0$ and $\poss(\action)\not\subseteq\state$. By
scheduling
\action,  the serial planner makes an assumption about the
observability of the precondition $\poss(\action)$.

Our approach always begins with a sequential session. Non-assumptive
actions from the trace are executed in sequence until the
applicability of the next scheduled action is uncertain. Here, that
action is denoted \switchAction.  A contingent session then begins,
which tailors sensory processing by the agent to determine whether the
assumptions made in the trace hold. For each {\em active} assumption
in the trace, we add a {\em disconfirm} action to the POMDP whose
execution is rewarding (resp. costly) if the assumption is false
(resp. true). We also add one {\em confirm} action that is rewarding
(resp. costly) to execute if $\poss(\action)$ is true
(resp. false). If execution of the contingent plan applies a
disconfirm action, then a new sequential session begins at the {\em
underlying} belief-state. If the confirm action is executed, the
sequential session is resumed, and
\switchAction\ is executed. Otherwise, control rests in the continual
session.
%%
Because contingent planning is only practical in relatively small
POMDPs, contingent sessions proceed in an abstract decision process
determined by the current trace. That abstract process is constructed
by first omitting all variables from the model that are not featured
in the trace, and by then iteratively refining that abstraction while
the resulting model is of a practicable size.

Finally, whether proceeding in a sequential or contingent session, our
continual planner maintains a factored representation of successive
{\em belief-states}, by performing belief revision according to
Eq~\ref{eq:revision}. Moreover, our internal representation of that
distribution corresponds more-or-less directly to an $(\pp{:init})$
declaration in DTPDDL. Our approach uses that underlying belief-state
as: (1) the source of determinisations for sequential planning, (2)
the basis for switching, and (3) the source for state abstraction in
contingent sessions.

@@

For sequential planning the operators from the POMDP model are
available with semantics that accommodate the above abstraction as
follows. In the deterministic model a proposition $\prop$ can be
thought to have a ternary interpretation at a state \state, as either
{\em true}, written $\prop\in\state$, {\em false}, written
$\prop\not\in\state$, or {\em unspecified}, with a slight abuse of
notation written $\prop\#\in\state$. For example, in $\pp{DORA}$
$(=(\pp{is-in}~\pp{box})\pp{office})\#\in\state_0$. For POMDP actions
$\action\in\actions$ in the deterministic model, if
$\prop \in \poss(\action)$, or if $\prop$ is the subject of a positive
or negative effect of $\action$, then $\action$ is not applicable
in \state\ if $\prop\#\in\state$. Indeed, first an assumptive action
must be applied to make a truth assignment to \prop, and then that
truth assignment could be modified by POMDP operators.

Addressing now the switching semantics of action execution. If
$\action\in\actions$ has a precondition that is not true with
probability $1$ in the underlying belief $\bstate$, then before it is
executed blindly, the system switches to a contingent session,
discussed in the following section. In detail, we halt the sequential
session at \action\ if $\exists
\state$ s.t. $\bstate(\state)>0$, 
$\prop\in\poss(\action)$ , and $\prop\not\in\state$. According to the
semantics of action execution there must be a $\state'$ s.t.
$\bstate(\state')>0$ and $\prop\in\state'$. Otherwise, \action\ must
not have been scheduled at \bstate.

then $\reward(\state,\action)=10$, and
otherwise if $\prop\not\in\state'$ then
$\reward(\state',\action)=-10$.
